Flowchart of Current Algorithm

### 2008-10-27 docs/flowchart/flowchart.tex
add some marginpars about volume discrepancy
Mike caught a sign error on volume discrepancy
fixed equations with volume discrepancy - units weren't right and the equations were written wrong

\item[plane-parallel:] We integrate
\begin{equation}
\frac{\partial w_0}{\partial r} =  \Sbar - \frac{1}{\gammabar p_0} \etarho g 
\enskip \label{eq:divw0} ,
\end{equation}
using the lagged $\psi^{\nmh}$ and the volume discrepancy correction with $f=0.3$,
\begin{equation}
\frac{\partial w_0^{\nph,\star}}{\partial r} =  \overline{S}^{\nph,\star} 
- \frac{1}{\gammabar^{n} p_0^{n}}
\left\{ \psi^{\nmh} - 
f \left[\frac{p_0^n - \overline{p(\rho,h,X_k)}^n}{\Delta t^n} \right]\right\}
\enskip .
\end{equation}

@@

The procedure to construct $\uadvone$ is described in detail in Appendix B of paper III.
We note that  $\uadvone$ satisfies the discrete versions of 
\begin{equation}
\int_{\Omega_H} \Ubt \cdotb \er \; dA = 0 \label{eq:udoterzero} \enskip ,
\end{equation}
(paper III; eq.~[21])
and
\begin{equation}
\nablab \cdotb (\beta_0 \Ubt )  = \beta_0 \left(S - \Sbar \right )\enskip ,
\label{eq:tildeconstraint}
\end{equation}
(paper III; eq.~[39]).  The discretization with the volume discrepancy correction is
\begin{equation}
\nablab \cdotb \left(\beta_0^n \uadvone\right) = 
\beta_0^n \left(S^{\nph,\star} - \overline{S}^{\nph,\star}\right)
+ \frac{f}{\gammabar^n p_0^n}
\left[\frac{p(\rho,h,X_k)^n - \overline{p(\rho,h,X_k)}^n}{\Delta t^n}\right] \enskip .
\end{equation}

@@

\item[plane-parallel:] We integrate equation (\ref{eq:divw0}) with the 
volume discrepancy correction,
\begin{equation}
\frac{\partial w_0^{\nph}}{\partial r} = 
\overline{S}^{\nph,\star} - \frac{1}{\gammabar^{\nph,\star} p_0^{\nph,\star}}
\left\{ \psi^{\nph,\star} - f \left[\frac{p_0^{\nph,\star} - \overline{p(\rho,h,X_k)}^{\nph,\star}}{\Delta t^n}\right]\right\}
\enskip ,
\end{equation}
\begin{equation}
\gammabar^{\nph,\star} = \frac{\gammabar^{n} + \gammabar^{n+1,\star}}{2}
\end{equation}
\begin{equation}
p_0^{\nph,\star} = \frac{p_0^{n} + p_0^{n+1,\star}}{2}
\end{equation}
\begin{equation}
\overline{p(\rho,h,X_k)}^{\nph,\star} = 
\frac{\overline{p(\rho,h,X_k)}^n + \overline{p(\rho,h,X_k)}^{n+1,\star}}{2}
\end{equation}

@@

The procedure to construct $\uadvtwo$ is described in detail in Appendix B of paper III
and is analogous to the procedure used in step 2, but with updated values
for $w_0$ and $\pizero.$  We note that $\uadvtwo$ satisfies the discrete versions of 
equations (\ref{eq:udoterzero}) and (\ref{eq:tildeconstraint}).   The discretization
with the volume discrepancy correction is
\begin{eqnarray}
\nablab \cdotb \left(\beta_0^{\nph,\star} \uadvtwo\right) &=& 
\beta_0^{\nph,\star} \left(S^{\nph} - \overline{S}^{\nph}\right) \nonumber \\
&& + \frac{f}{\gammabar^{\nph,\star} p_0^{\nph,\star}}
\left[\frac{p(\rho,h,X_k)^{\nph,\star} - \overline{p(\rho,h,X_k)}^{\nph,\star}}{\Delta t^n}\right]
\enskip , \nonumber \\
\end{eqnarray}
\begin{equation}
\beta_0^{\nph,\star} = \frac{ \beta_0^n +  \beta_0^{n+1,\star} }{2} \enskip .
\end{equation}

@@

Finally, we impose the divergence constraint from equation (\ref{eq:tildeconstraint}) 
including the volume discrepancy correction,
\begin{equation}
\nablab \cdotb \left(\beta_0^{\nph} \Ubt^{n+1} \right)  = \beta_0^{\nph} \left(S^{n+1} - \overline{S}^{n+1} \right)
+ \frac{f}{\gammabar^{n+1} p_0^{n+1}}
\left[\frac{p(\rho,h,X_k)^{n+1} - \overline{p(\rho,h,X_k)}^{n+1}}{\Delta t^n}\right]
\enskip,
\end{equation}
by defining ${\bf V} = \Ubt^{n+1,\star} + ( \dt / \rho^\nph ) \; \mathbf{G} \pi^\nmh$ 
and solving
\begin{equation}
 L_\beta^\rho \phi =
   D \left ( \beta_0^{\nph} {\bf V} \right) - \beta_0^{\nph} \left(S^{n+1}-\overline{S}^{n+1} \right)
- \frac{f}{\gammabar^{n+1} p_0^{n+1}}
\left[\frac{p(\rho,h,X_k)^{n+1} - \overline{p(\rho,h,X_k)}^{n+1}}{\Delta t^n}\right] \enskip,
\end{equation}
for nodal values of $\phi$, where
$\beta_0^\nph = ( \beta_0^n + \beta_0^{n+1} )/2$ and
$L_\beta^\rho$ is the standard bilinear
finite element approximation to $\nablab \cdotb ({\beta_0^\nph}/{\rho^\nph}) \nablab.$
\MarginPar{Why are $\beta$ and $\rho$ time centered here instead using the new time value?  Does this have something to do with how $\phi$ relates to a lagged pressure?}
In this step, $D$ is a discrete second-order operator that approximates the 
divergence at nodes from cell-centered data and satisfies
$\mathbf{G} = -D^T.$ 
(See (cite almgrenBellSzymczak:1996) for a detailed discussion of this
approximate projection; see (cite almgren:bell:crutchfield) for a discussion
of this particular form of the projection operand.)  
We solve the linear system of equation
\begin{equation}
 L_\beta^\rho \phi =
   D \left ( \beta_0^{\nph} {\bf V} \right) - \beta_0^{\nph} \left(S^{n+1}-\overline{S}^{n+1} \right)\enskip,
\label{eq:nodal_solve}
\end{equation}
(paper III; eq.~[77]) using multigrid V-cycles with Gauss-Seidel relaxation.

>>>>

\item[plane-parallel:] We integrate
\begin{equation}
\frac{\partial w_0}{\partial r} =  \Sbar - \frac{1}{\gammabar p_0} \etarho g 
\enskip \label{eq:divw0} ,
\end{equation}
using the lagged $\psi^{\nmh}$ and the volume discrepancy correction with $f=0.3$,
\begin{equation}
\frac{\partial w_0^{\nph,\star}}{\partial r} =  \overline{S}^{\nph,\star} 
- \frac{1}{\gammabar^{n} p_0^{n}}
\left\{ \psi^{\nmh} + 
f \left[\frac{p_0^n - \overline{p(\rho,h,X_k)}^n}{\Delta t^n} \right]\right\}
\enskip .
\end{equation}

@@

The procedure to construct $\uadvone$ is described in detail in Appendix B of paper III.
We note that  $\uadvone$ satisfies the discrete versions of 
\begin{equation}
\int_{\Omega_H} \Ubt \cdotb \er \; dA = 0 \label{eq:udoterzero} \enskip ,
\end{equation}
(paper III; eq.~[21])
and
\begin{equation}
\nablab \cdotb (\beta_0 \Ubt )  = \beta_0 \left(S - \Sbar \right )\enskip ,
\label{eq:tildeconstraint}
\end{equation}
(paper III; eq.~[39]).  The discretization with the volume discrepancy correction is
\begin{equation}
\nablab \cdotb \left(\beta_0^n \uadvone\right) = 
\beta_0^n \left\{ \left(S^{\nph,\star} - \overline{S}^{\nph,\star}\right)
+ \frac{f}{\gammabar^n p_0^n}
\left[\frac{p(\rho,h,X_k)^n - \overline{p(\rho,h,X_k)}^n}{\Delta t^n}\right]\right\} \enskip .
\end{equation}

@@

\item[plane-parallel:] We integrate equation (\ref{eq:divw0}) with the 
volume discrepancy correction,
\begin{equation}
\frac{\partial w_0^{\nph}}{\partial r} = 
\overline{S}^{\nph,\star} - \frac{1}{\gammabar^{\nph,\star} p_0^{\nph,\star}}
\left\{ \psi^{\nph,\star} + f \left[\frac{p_0^{\nph,\star} - \overline{p(\rho,h,X_k)}^{\nph,\star}}{\Delta t^n}\right]\right\}
\enskip ,
\end{equation}
\begin{equation}
\gammabar^{\nph,\star} = \frac{\gammabar^{n} + \gammabar^{n+1,\star}}{2}
\end{equation}
\begin{equation}
p_0^{\nph,\star} = \frac{p_0^{n} + p_0^{n+1,\star}}{2}
\end{equation}
\begin{equation}
\overline{p(\rho,h,X_k)}^{\nph,\star} = 
\frac{\overline{p(\rho,h,X_k)}^n + \overline{p(\rho,h,X_k)}^{n+1,\star}}{2}
\end{equation}

@@

The procedure to construct $\uadvtwo$ is described in detail in Appendix B of paper III
and is analogous to the procedure used in step 2, but with updated values
for $w_0$ and $\pizero.$  We note that $\uadvtwo$ satisfies the discrete versions of 
equations (\ref{eq:udoterzero}) and (\ref{eq:tildeconstraint}).   The discretization
with the volume discrepancy correction is
\begin{equation}
\nablab \cdotb \left(\beta_0^{\nph,\star} \uadvtwo\right) =
\beta_0^{\nph,\star}\left\{\left(S^{\nph} - \overline{S}^{\nph}\right)
 + \frac{f}{\gammabar^{\nph,\star} p_0^{\nph,\star}}
\left[\frac{p(\rho,h,X_k)^{\nph,\star} - \overline{p(\rho,h,X_k)}^{\nph,\star}}{\Delta t^n}\right]\right\} \enskip ,
\end{equation}
\begin{equation}
\beta_0^{\nph,\star} = \frac{ \beta_0^n +  \beta_0^{n+1,\star} }{2} \enskip .
\end{equation}

@@

Finally, we impose the divergence constraint from equation (\ref{eq:tildeconstraint}) 
including the volume discrepancy correction,
\begin{equation}
\nablab \cdotb \left(\beta_0^{\nph} \Ubt^{n+1} \right)  = \beta_0^{\nph}\left\{  \left(S^{n+1} - \overline{S}^{n+1} \right)
+ \frac{f}{\gammabar^{n+1} p_0^{n+1}}
\left[\frac{p(\rho,h,X_k)^{n+1} - \overline{p(\rho,h,X_k)}^{n+1}}{\Delta t^n}\right]\right\}
\enskip,
\end{equation}
by defining ${\bf V} = \Ubt^{n+1,\star} + ( \dt / \rho^\nph ) \; \mathbf{G} \pi^\nmh$ 
and solving
\begin{equation}
 L_\beta^\rho \phi =
   D \left ( \beta_0^{\nph} {\bf V} \right) - \beta_0^{\nph}\left\{ \left(S^{n+1}-\overline{S}^{n+1} \right)
- \frac{f}{\gammabar^{n+1} p_0^{n+1}}
\left[\frac{p(\rho,h,X_k)^{n+1} - \overline{p(\rho,h,X_k)}^{n+1}}{\Delta t^n}\right]\right\} \enskip,
\end{equation}
for nodal values of $\phi$, where
$\beta_0^\nph = ( \beta_0^n + \beta_0^{n+1} )/2$ and
$L_\beta^\rho$ is the standard bilinear
finite element approximation to $\nablab \cdotb ({\beta_0^\nph}/{\rho^\nph}) \nablab.$
\MarginPar{Why are $\beta$ and $\rho$ time centered here instead using the new time value?  Does this have something to do with how $\phi$ relates to a lagged pressure?}
In this step, $D$ is a discrete second-order operator that approximates the 
divergence at nodes from cell-centered data and satisfies
$\mathbf{G} = -D^T.$ 
(See (cite almgrenBellSzymczak:1996) for a detailed discussion of this
approximate projection; see (cite almgren:bell:crutchfield) for a discussion
of this particular form of the projection operand.)  
We solve this linear system of equations using multigrid V-cycles with 
Gauss-Seidel relaxation.
